% Admissible Non-Injective Transitions as the Primitive of Physical Description
% Cosmochrony companion paper M
% Provides the axiomatic foundation cited by Q1, Bell, Beau2026l, and the white paper.

\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumerate}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}

\bibliographystyle{plain}
\usepackage{doi}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{principle}[theorem]{Principle}

\title{Admissible Non-Injective Transitions\\
as the Primitive of Physical Description}

\author{J\'er\^ome Beau\\
\small Independent Researcher\\
\small \texttt{jerome.beau@cosmochrony.org}}

\date{2026}

\begin{document}

  \maketitle

  \begin{abstract}
    Quantum mechanics is the effective theory of physically real unresolved
    admissible structures and their resolution.
    This paper establishes the minimal axiomatic foundation from which this
    claim follows: physical structure is derived from a single primitive,
    the local structure of admissible transitions between observable states.
    Four axioms govern this structure: local projective admissibility (A1),
    structural non-injectivity (A2), admissibility as non-premature selection (A3),
    and projection locking (A4).
    From A1 and A2 alone, we derive irreversibility, the arrow of time, and the
    existence of structural fluctuations, without any additional postulate.
    From A3, we derive the proto-state as a physically real unresolved configuration
    shared between two successive states, and show that admissibility forces the
    retention of phase coherence.
    We further prove that A3 forbids any admissible factorisation of the fibre,
    which forces a non-trivial commutator between conjugate generators;
    combined with minimality (A1) and Born--Infeld parity, this identifies the
    symmetry group of the admissible fibre as $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$
    for a prime $q$, and the fibre itself as the Weil representation $V_\rho$ —
    closing the identification $F_n \simeq V_\rho$ as a theorem rather than a remark.
    From A4, we derive the discrete character of quantum transitions as a consequence
    of the interaction between continuous Born--Infeld saturation and the discrete
    shell structure of the observable space.
    This paper provides the axiomatic foundation cited by the companion
    papers~\cite{Beau2026q1,Beau2026b,Beau2026l} and serves as the formal basis for the derivations
    carried out therein.
  \end{abstract}

  \paragraph{Keywords.}
    Non-injective projection; Emergence; Projective structure; Admissibility;
    Born--Infeld constraint; Information loss;
    Projection entropy; Quantum emergence; Phase coherence; SU(2) structure;
    Gauge emergence; Fiber structure;
    Projective temporal ordering; Irreversibility.

    \clearpage
    \tableofcontents

    \clearpage

  \section{Introduction}
  \label{sec:intro}

  The Cosmochrony programme~\cite{Cosmochrony} aims to derive the structure of
  physical description from a minimal relational foundation, without postulating
  a background spacetime, a Hilbert space, or an external substrate.
  The only primitive is the local structure of admissible transitions between
  observable states.

  The central question driving the framework is:

  \begin{quote}
    Given an observable state $O_{n-1}$, what states are accessible from it?
  \end{quote}

  The answer to this question — encoded in the admissibility structure — is
  sufficient to derive irreversibility, temporal order, wave-like propagation,
  phase coherence, and the discrete character of quantum transitions.
  No substrate, no background, and no quantum postulate is required as input.

  \subsection*{Relation to companion papers}

    This paper provides the axiomatic foundation for a chain of companion results.
    The companion paper~\cite{Beau2026b} establishes that Bell-type factorisation
    fails generically under non-injective projection (A2), without invoking
    nonlocal dynamics.
    The companion paper~\cite{Beau2026l} proves that non-injectivity is a structural
    necessity of any genuinely emergent description.
    The companion paper~\cite{Beau2026fe} develops the information-theoretic
    architecture of the resulting projection loss:
    it establishes a monotone fibre-erasure functional $I(c;\sigma(\ell))$
    non-increasing under admissible coarse-graining,
    connecting the ENI projection entropy to the BFS capacity hierarchy
    of the spectral programme.
    The companion paper~\cite{Beau2026q1} derives phase coherence, the singlet correlator,
    the Tsirelson bound, and the Born rule as structural consequences of
    admissibility — using Axioms A1--A4 as stated here, combined with the concrete
    spectral results of the O-series.
    The companion papers~\cite{Beau2026entnote,Beau2026ent1,Beau2026ent2} locate entanglement
    quantitatively within the spectral admissibility cascade:
    the minimal admissible pair sector carries the spin-$\tfrac12$ singlet with entropy
    $\log 2$ (the Bell--Tsirelson quantity), while the full residual fibre-level rank
    gives a finer observable $S_{\mathrm{ent}}(n)=\log r_{\mathrm{pair}}(n)$.
    The white paper~\cite{Cosmochrony} develops the full phenomenological programme;
    the present axiomatic reformulation is a foundational companion to it, not yet
    integrated into its main text.

  \subsection*{Organisation}

    Section~\ref{sec:axioms} states the four axioms.
    Section~\ref{sec:primitive} clarifies the role of $\Pi$ and the index $n$.
    Section~\ref{sec:ni-consequences} derives the structural consequences of
    non-injectivity: irreversibility, novelty, temporal order, and the
    projective incompleteness principle.
    Section~\ref{sec:admissibility} derives the proto-state, phase, coherence,
    and the Heisenberg structure of the admissible fibre from the admissibility
    principle.
    Section~\ref{sec:locking} derives projection locking and the discrete character
    of quantum transitions.
    Section~\ref{sec:entanglement} treats entanglement and Bell non-applicability.
    Section~\ref{sec:summary} summarises what is derived from what, and identifies
    the open directions.

    \clearpage

  \section{The Four Axioms}
  \label{sec:axioms}

  We state the four axioms from which the framework proceeds.
  They are not independent postulates chosen for convenience — each captures
  a structural feature that is either necessary for genuine emergence~\cite{Beau2026l}
  or forced by the requirement of bounded effective descriptions~\cite{Beau2026c}.

  \begin{definition}[Admissible transition]
    \label{def:admissible-transition}
    A transition from observable state $O_{n-1}$ to observable state $O_n$ is
    \emph{admissible} if it satisfies Axioms A1--A4 below.
  \end{definition}

  \begin{remark}[Admissibility is relational, not absolute]
    \label{rem:admissibility-relational}
    Admissibility is never an intrinsic predicate of an observable state $O_n$ taken in
    isolation.
    It is a relational constraint on projective transitions: a state $O_n$ is admissible
    only relative to a preceding state $O_{n-1}$, whose accumulated constraints determine
    the local admissible neighbourhood.
    We write $O_n \in \mathcal{A}(O_{n-1})$ to make this explicit, where
    \begin{equation}
      \label{eq:admissible-neighbourhood}
      \mathcal{A}(O_{n-1}) \;:=\;
      \bigl\{\, O_n \mid \text{the transition } O_{n-1} \to O_n \text{ satisfies A1--A4} \,\bigr\}
    \end{equation}
    is the \emph{admissible neighbourhood from $O_{n-1}$}.
    The phrasing ``$O_n$ is admissible'' should always be read as shorthand for
    $O_n \in \mathcal{A}(O_{n-1})$: admissibility is always \emph{from-here} admissibility,
    conditioned on the already-realized projective position.
    There is no fixed space of possible observable states: there is a field of local
    admissible neighbourhoods, each determined by the point already reached.
  \end{remark}

  \begin{description}
    \item[\textbf{A1 — Local projective admissibility.}]
    For any observable state $O_{n-1}$, there exists a non-empty set $F_n$ of
    admissible successor directions: configurations of $O_{n-1}$ compatible with
    a transition to a stable next state.
    The set $F_n$ is defined by internal structural constraints
    (Born--Infeld bound, spectral envelope, symmetry requirements)
    and does not presuppose an ambient configuration space.

    \item[\textbf{A2 — Structural non-injectivity.}]
    Distinct admissible directions from $O_{n-1}$ may lead to the same resolved
    successor state $O_n$.
    The map $\Pi_n$ encoding this is generically non-injective.

    \item[\textbf{A3 — Admissibility as non-premature selection.}]
    An admissible transition preserves the full multiplicity of open directions
    until projection locking occurs.
    It does not select among them prematurely — that is, before the
    Born--Infeld saturation threshold is reached~\cite{Beau2026c}.

    \item[\textbf{A4 — Projection locking.}]
    Resolution occurs only when continuous Born--Infeld saturation meets the
    discrete shell support of the observable space.
    The transition from admissible to admitted is structurally discrete.
  \end{description}

  \begin{remark}[Independence of the axioms]
  A1 asserts existence of admissible directions.
  A2 asserts their generic non-injectivity.
  A3 constrains which transitions are admissible.
  A4 governs when resolution occurs.
  The four axioms are logically independent: none follows from the others.
  Their joint necessity is argued in Section~\ref{sec:primitive} and in the
  companion papers~\cite{Beau2026l,Beau2026c}.
  \end{remark}

  \begin{remark}[Status of Born--Infeld in A3 and A4]
  The Born--Infeld saturation threshold appearing in A3 and A4 is not an
  additional postulate: it is the unique bounded completion of the effective
  action compatible with bounded projective flux, established in~\cite{Beau2026c}.
  A3 and A4 therefore do not introduce new structure beyond A1--A2 and the
  Born--Infeld companion.
  \end{remark}

  \begin{remark}[What projection locking selects]
    \label{rem:a4-selects}
    A4 resolves the transition to a unique observable state~$O_n$ in the image of~$\Pi_n$.
    It does not select a unique underlying configuration in~$F_n$.

    The multiplicity of admissible configurations in the fibre $F_n = \Pi_n^{-1}(O_n)$ is
    not eliminated by projection locking: distinct configurations $f, f' \in F_n$ with
    $\Pi_n(f) = \Pi_n(f') = O_n$ remain structurally present after locking, but become
    observationally inaccessible.
    The non-injectivity of~$\Pi_n$ therefore persists at every resolution step.

    This distinction --- observable uniqueness without configurational uniqueness --- prevents
    Bell-type factorisation~\cite{Beau2026b} and forces the coherent, non-factorisable fibre
    structure derived in Section~\ref{subsec:heisenberg}.
    More generally, distinct interactions $I_1 \neq I_2$ induce distinct admissible
    sub-fibres $F_n^{(I_1)} \neq F_n^{(I_2)}$ and hence distinct observable outcomes,
    without any element of $F_n$ having been selected.
    This is the structural origin of contextuality: measurement context determines
    which sub-fibre is accessed, not which configuration is realised.
    This contextuality is \emph{synchronic}: it concerns the choice of sub-fibre at a
    fixed transition step $n$.
    The admissible neighbourhood $\mathcal{A}(O_{n-1})$ (Remark~\ref{rem:admissibility-relational})
    introduces a second, \emph{diachronic} form of contextuality: the set of accessible
    observable states at step $n$ is itself conditioned on the previously realized state
    $O_{n-1}$.
    The context of a projection is therefore not merely the measurement algebra chosen at
    step $n$, but also the already-realized projective position from which the next state
    may be reached.
    Standard contextuality frameworks (Kochen--Specker, sheaf-theoretic) capture synchronic
    contextuality; the present framework extends this to diachronic contextuality, in which
    the history of realized projections is constitutive of what is admissible next.
  \end{remark}

  \begin{remark}[Projective Observability Principle]
    \label{rem:projective-observability}
    Axioms A1--A4 jointly imply the following criterion, which identifies what is
    physical in the framework.

    \emph{Physical observables coincide with the invariants of the projection~$\Pi_n$:
    any quantity that depends on the choice of pre-image within~$\Pi_n^{-1}(O_n)$
      is unphysical by construction.}

    This is not an additional axiom but a consequence of A2 (which forces non-trivial
    fibres) and A4 (which identifies $O_n$ with $\mathrm{Im}\,\Pi_n$ exactly):
    since $\Pi_n$ is the only structure that relates the configurational level to the
    observable level, any quantity invisible to $\Pi_n$ is invisible to physics.
    What the projection erases --- intra-fibre distinctions --- is unphysical;
    what it preserves --- the invariants of the fibre structure --- constitutes the
    physical content of the theory.
    In particular, the symmetry group of a fibre $\Pi_n^{-1}(O_n)$ is physical precisely
    because it acts within the fibre while leaving $O_n$ unchanged: this is the structural
    origin of gauge symmetry, developed in the companion papers~\cite{Beau2026q6a,Beau2026a35}.
    A detailed formulation in terms of the five-level hierarchy is given in
    Remark~\ref{rem:observable-space}.
  \end{remark}

  \clearpage
  \section{The Primitive and the Role of $\Pi$}
  \label{sec:primitive}

  \subsection{The primitive}

    The only primitive of this framework is the pair $(O_{n-1}, F_n)$: an observable
    state and the set of admissible directions it admits.
    There is no external substrate, no hidden carrier, no background geometry.
    The fibre $F_n$ is not defined in an abstract ambient space — it is induced by
    $O_{n-1}$ itself through the admissibility constraints of A1--A3.

    The resulting evolution is recursive: once $O_n$ is realized, it immediately becomes
    the new departure point, generating the next admissible neighbourhood $\mathcal{A}(O_n)$
    from which $O_{n+1}$ may be reached:
    \begin{equation}
      \label{eq:recursive-chain}
      O_{n-1}
      \;\longrightarrow\; O_n
      \;\longrightarrow\; \mathcal{A}(O_n)
      \;\longrightarrow\; O_{n+1}.
    \end{equation}
    The realized state is simultaneously a destination and the generator of the next set of
    constraints.
    No external agent resets the context between steps: the admissible neighbourhood at
    each stage is determined entirely by the state just reached.

  \subsection{The projection $\Pi_n$ as descriptive abstraction}

    The notation $\Pi_n$ is a compact encoding of the local admissibility structure.
    It encodes three things simultaneously:
    \begin{enumerate}[(i)]
\item the admissible directions from $O_{n-1}$ — the fibre $F_n$ (A1, A2);
\item the Born--Infeld saturation bound — the limit on accessible amplitudes (A3);
\item projection locking — the constraint that resolution falls on an available
shell (A4).
    \end{enumerate}

    $\Pi_n$ is not a physical entity and not a dynamical map.
    It is a structural relation encoding the admissible identification between
    distinguishable configurations (elements of $F_n$) and indistinguishable
    observable outcomes (elements of $O_n$).
    Saying ``$\Pi_n$ is non-injective'' means that multiple directions from
    $O_{n-1}$ can lead to the same $O_n$ — a fact about the admissible transition
    structure, not a property of an external map.
    Non-injectivity does not originate from an underlying substrate, but from
    the structure of admissible configurations and their interaction-dependent
    realisation map.

    \begin{remark}[The threefold role of $\Pi_n$]
      \label{rem:pi-threefold}
      The structural consequences established below in Sections~\ref{sec:ni-consequences}
      and~\ref{subsec:projective-incompleteness} reveal that $\Pi_n$ plays three
      distinct roles simultaneously, each irreducible to the others.
      \begin{enumerate}[(i)]
        \item \emph{Admissibility filter.}
        $\Pi_n$ selects, from the structurally compatible continuations
        $\Omega_n^{(c)}$, those directions satisfying the Born--Infeld and spectral
        constraints of A1--A3.
        What it retains constitutes the observable $O_n$; what it discards is the
        non-trivial kernel whose existence is guaranteed by
        Corollary~\ref{cor:projective-incompleteness}.
        \item \emph{Generator of temporal order.}
        The oriented sequence $\Pi_0, \Pi_1, \Pi_2, \ldots$ carries the partial order
        established in Theorem~\ref{thm:temporal-order}.
        Each $\Pi_n$ is a node in the admissibility graph
        (Remark~\ref{rem:state-machine}); the index $n$ acquires its temporal meaning
        from this orientation, not from an external parameter.
        The dynamics of the framework is the oriented topology of this graph — not
        a law superimposed on a pre-existing time.
        \item \emph{Revelation, not creation.}
        $\Pi_n$ does not produce new structure in $\chi$.
        It selects, from the admissible spectral modes of the Weil representation
        already present in $\Omega_n^{(c)}$, those that are observationally
        accessible at depth $n$.
        The BFS depth controls how much of the structure of $\chi$ has been
        revealed; increasing $n$ does not modify $\chi$ but resolves finer spectral
        content — as a deeper lens resolves detail that was always present in the
        object, not detail that the lens creates.
      \end{enumerate}
      These three roles are inseparable: the structural constraints A1--A3 and
      Born--Infeld parity that force the admissibility filter simultaneously
      generate the temporal order via Theorem~\ref{thm:temporal-order} and
      guarantee the incompleteness of every resolution via
      Corollary~\ref{cor:projective-incompleteness}.
      The non-commutativity $[X, \sigma(X)] \neq 0$ at the root of the
      incompleteness (Theorem~\ref{thm:heisenberg}) is therefore not merely a
      technical property of the fibre: it is the algebraic source of all three
      roles at once.
    \end{remark}

  \subsection{The index $n$}

    The index $n$ does not represent a pre-existing time parameter.
    It counts transitions: the number of times a state has been succeeded by
    another under the admissibility constraints.
    Its temporal meaning is derived in Section~\ref{sec:ni-consequences} —
    it is not assumed.

  \subsection{Where non-injectivity lives}
    \label{subsec:ni-locus}

    A natural question arises: where does the non-injectivity of $\Pi_n$ come from,
    given that no underlying substrate is postulated?
    The answer is precise.

    Non-injectivity is a property of the interaction-dependent realisation map:
    the map from the space $F_n$ of structurally compatible configurations to the space
    $O_n$ of observable outcomes.
    When two distinct admissible configurations $f, f' \in F_n$ give rise to
    the same observable outcome $o \in O_n$, it is not because they are
    ``the same at a deeper level'' — it is because the interaction that
    realises the transition cannot distinguish them within $O_n$.
    The limit on distinguishability is set by the Born--Infeld capacity
    constraint (A3): the observable space $O_n$ has finite projective capacity,
    and configurations that differ only below this threshold are identified.

    Three properties of this realisation map are essential:
    \begin{enumerate}[(i)]
\item \emph{Interaction-dependence}: $\Pi_n$ encodes the interaction;
different interactions define different realisation maps and different
fibres $F_n$.
\item \emph{Internal definition of admissibles}: $F_n$ is defined by
internal constraints (BI bound, spectral envelope, symmetry) — not by
reference to an ambient substrate.
\item \emph{Minimal structure of $F_n$}: the admissibility constraints
force $F_n$ to carry a specific algebraic structure — SU(2) in the
spin-$\tfrac{1}{2}$ sector (O23) — which is derived from the constraints,
not assumed.
    \end{enumerate}

    The non-injectivity is therefore neither a property of a hidden substrate
    nor an arbitrary assumption: it is a structural consequence of finite
    projective capacity applied to an internally defined space of admissible
    configurations.

    \begin{remark}[The admissible fibre as equivalence class under projection]
      \label{rem:shadow-analogy}
      A useful image for the non-injectivity of~$\Pi_n$ is that of shadows cast by a fixed
      projection setup.
      A given shadow does not correspond to a single object:
      it corresponds to the set of all objects whose projection yields that shadow ---
      the admissible fibre $F_n = \Pi_n^{-1}(O_n)$.

      The directionality is essential.
      The fibre is not the collection of all shadows produced by a given object under varying
      projections (one configuration, many images).
      It is the collection of all configurations that produce the same observable under a fixed,
      non-injective~$\Pi_n$ (many configurations, one observable).
      This many-to-one structure is the source of irreversibility
      (Section~\ref{sec:ni-consequences}) and of the non-factorisability of the proto-state
      (Section~\ref{subsec:heisenberg}).
    \end{remark}

    \begin{remark}[Connection to the O-series: $F_n$ and $V_\rho$]
      \label{rem:fn-vrho}
      In the concrete spectral realisation of the O-series, the admissible fibre
      $F_n$ is realised as a spectral sector $V_\rho$ of the Heisenberg Cayley
      graph $\mathrm{Cay}(\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z}), S_q)$.
      The sectors $V_\rho$ are not pre-existing configuration spaces: they are
      the structures induced by the admissibility constraints — the Born--Infeld
      bound, the spectral envelope, and the symmetry requirements of A1.
      In this sense, $V_\rho$ is the concrete realisation of $F_n$, not an
      ambient space in which $F_n$ is embedded:
      \begin{equation}
        F_n \;\simeq\; V_\rho.
      \end{equation}
      This identification, stated here as a structural connection to the O-series,
      is established as a theorem in Section~\ref{subsec:heisenberg}:
      the axioms A1--A3 and Born--Infeld parity together force the symmetry group
      of $F_n$ to be $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$, and the
      Stone--von Neumann theorem then identifies $F_n \simeq V_\rho$ uniquely.
    \end{remark}

    \begin{remark}[The two levels of $\Pi_n$: kernel and image (O24)]
      \label{rem:pi-two-levels}
      The non-injective projection $\Pi_n$ operates at two structurally distinct
      levels, established in O24~\cite{Beau2026a28}:
      \begin{enumerate}[(i)]
  \item \emph{Kernel level} ($\ker \Pi_n$): governs microscopic multiplicity
  and fluctuation amplitude.
  Multiple configurations in $\Omega_n^{(c)}$ may project to the same
  observable outcome; their differences are invisible to admissible
  observables and appear as projection residuals.
  The dependence on $O_{n-1}$ — via the baseline $V_{n-1}^{(c)}$ — enters
  at this level: it determines which directions in $\Omega_n^{(c)}$ are
  new relative to the already realized subspace.
  \item \emph{Image level} ($\mathrm{Im}\,\Pi_n \cap \mathrm{Ntrl}$):
  governs the observable structure.
  Its real rank is rigidly fixed to 3 by the quaternionic maximality of
  $\mathrm{Im}\,\mathbb{H}$ (O23, Theorem~3.2) and is independent of the
  cardinality of $\ker \Pi_n$.
      \end{enumerate}
      The Verticality Lemma of O24 (Lemma~3.1) establishes that every
      Born--Infeld-admissible symmetry acts vertically: it maps each fibre
      $\Pi_n^{-1}(o)$ to itself and does not increase the rank of the image.
      The structural consequence is:
      \begin{quote}
        \emph{Non-injectivity affects degeneracy, not structure.}
      \end{quote}
      Enlarging $\ker \Pi_n$ — admitting more configurations in each fibre —
      increases the amplitude of quantum fluctuations but does not alter the
      observable rank, the cascade exponent $\delta_{\mathrm{pair}}$, or any
      other structural invariant of the admissible sector.
      The two levels of $\Pi_n$ are therefore decoupled: kernel structure is
      physically relevant for fluctuations; image structure is physically
      relevant for observables.
    \end{remark}

    \begin{remark}[The observable space principle: $\mathcal{O} = \mathrm{Im}\,\Pi$]
      \label{rem:observable-space}
      A direct consequence of the five-level hierarchy is that admissible
      observables are functions on
      \begin{equation}
        \mathcal{O} \;:=\; \mathrm{Im}\,\Pi
        \;=\; \bigsqcup_{n} O_n,
      \end{equation}
      not on the pre-image space~$\Omega_n^{(c)}$.
      Any structural difference between two configurations
      $f, f' \in \Pi_n^{-1}(O_n)$ that share the same image under~$\Pi_n$
      is invisible to every admissible observable by definition:
      it is a difference \emph{within} the fibre, which the projection
      identifies as observationally indistinguishable.
      This is not an additional assumption but an immediate consequence of
      Definition~\ref{def:omega-n} and Level~5:
      $O_n$ is defined as $\mathrm{Im}\,\Pi_n$, and admissibility is
      a property of $\Pi_n$ acting on $\Omega_n^{(c)}$, not of
      $\Omega_n^{(c)}$ itself.
      Intra-fibre variability is therefore structurally unobservable; it
      constitutes the degree of freedom whose projected manifestation appears
      as quantum indeterminacy in the effective description.
      In particular, any admissible quantity that depends on the choice of
      representative in $\Pi_n^{-1}(O_n)$ must agree for all representatives
      in the same fibre class: this is the structural basis for the uniformity
      of admissible spectral invariants across the fibre.
      The information-theoretic quantification of this erasure ---
      as a monotone non-increasing functional under admissible coarse-graining
      --- is developed in~\cite{Beau2026fe}.
    \end{remark}


    \label{subsec:five-levels}

    The abstract structure of Section~\ref{sec:primitive} has a concrete realisation
    in the O-series that answers precisely where $\Omega_n$ lives and how it is
    constructed.
    The realisation proceeds through five levels, each induced by the previous one.

    \paragraph{Level 1 — The relational group $G_q$.}
      The combinatorial substrate of the BFS exploration is the discrete Heisenberg
      group
      \begin{equation}
        G_q \;=\; \mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z}),
        \qquad |G_q| = q^3,
      \end{equation}
      equipped with the standard symmetric generating set $S_q = \{X^{\pm 1}, Y^{\pm 1}\}$.
      $G_q$ is not a physical space: it is a relational structure, encoding how
      configurations relate to one another via the generators.
      The BFS explores $G_q$ in shells $S_n = B_n \setminus B_{n-1}$, where
      $B_n = \{g \in G_q : d(e,g) \leq n\}$.
      The shell structure $\{S_n\}$ is the concrete realisation of the
      transition-counting index $n$ of Section~\ref{sec:primitive}.

    \paragraph{Level 2 — The Weil representation $\rho_c$ and its ambient space.}
      For each non-zero character $c \in (\mathbb{Z}/q\mathbb{Z})^\times$, the
      discrete Heisenberg group $G_q$ admits an irreducible $q$-dimensional Weil
      representation~\cite{Beau2026a16}:
      \begin{equation}
        \rho_c : G_q \;\longrightarrow\; \mathrm{GL}(\mathbb{C}^q),
        \qquad
        (\rho_c(a,b,\gamma)\,f)(x) \;=\;
        e^{2\pi i c(\gamma + bx)/q}\, f(x-a),
      \end{equation}
      acting on $L^2(\mathbb{Z}/q\mathbb{Z}) \cong \mathbb{C}^q$.
      The ambient space $\mathbb{C}^q$ is not a physical configuration space
      presupposed by the theory: it is the natural carrier of the irreducible
      representation of $G_q$ of central character $c$.
      It is induced by $G_q$ and $c$, not by any external assumption.

    \paragraph{Level 3 — The orbit span $\Omega_n$ as emerging subspace.}
      Given the initial vector $v_0 \in \mathbb{C}^q$ and the Weil action $\rho_c$,
      the orbit span at BFS depth $n$ is:
      \begin{equation}
        \label{eq:omega-n}
        \Omega_n^{(c)}
        \;:=\;
        \mathrm{span}\bigl\{\rho_c(g)\,v_0 : g \in B_n\bigr\}
        \;\subseteq\; \mathbb{C}^q.
      \end{equation}
      By construction, $\Omega_n^{(c)}$ is a minimal extension of the previously
      realized subspace:
      \begin{equation}
        \label{eq:omega-n-inclusion}
        V_{n-1}^{(c)} \;\subseteq\; \Omega_n^{(c)} \;\subseteq\; \mathbb{C}^q,
      \end{equation}
      where $V_{n-1}^{(c)} = \mathrm{GS}(B_{n-1})$ is the Gram--Schmidt span
      accumulated over the preceding shells.
      $\Omega_n^{(c)}$ is not defined by $O_{n-1}$, but is constrained by the
      already realized subspace $V_{n-1}^{(c)}$: the directions in
      $\Omega_n^{(c)} \setminus V_{n-1}^{(c)}$ are the candidates for new
      observable structure at depth $n$.

      \begin{definition}[Structurally compatible continuations]
        \label{def:omega-n}
        $\Omega_n^{(c)}$ is the space of \emph{structurally compatible continuations}
        from $O_{n-1}$: it contains every direction reachable by the relational
        structure of $G_q$ via $\rho_c$, without any admissibility filter applied.
        Admissibility is \emph{not} a property of $\Omega_n^{(c)}$;
        it is a property of $\Pi_n$ acting on $\Omega_n^{(c)}$.
      \end{definition}

      Three properties follow immediately:
      \begin{enumerate}[(i)]
\item \emph{Emergence}: $\Omega_n^{(c)}$ is not given a priori — it is built
incrementally as the BFS explores $G_q$.
\item \emph{Indirect conditioning on $O_{n-1}$}:
$\Omega_n^{(c)}$ is defined by the group action alone — the definition
\eqref{eq:omega-n} refers only to $B_n$ and $\rho_c$.
The dependence on $O_{n-1}$ is indirect: $O_{n-1}$ determines the
Gram--Schmidt span $\mathrm{GS}_{n-1}$ already accumulated over $B_{n-1}$,
which acts as the baseline against which new directions in $\Omega_n^{(c)}$
are measured.
This baseline is used by $\Pi_n$ (not by $\Omega_n^{(c)}$ itself) to
identify genuinely new directions, counted by $\sigma_c(n) = \delta r_n / |S_n|$~\cite{Beau2026a16}.
In summary: $\Omega_n^{(c)}$ is determined by $(G_q, \rho_c, B_n)$;
the role of $O_{n-1}$ enters only through $\Pi_n$.
\item \emph{Pre-admissibility character}:
$\Omega_n^{(c)}$ contains all directions generated by the group action
of $G_q$ up to depth $n$ — including directions that $\Pi_n$ will
subsequently reject as inadmissible.
The admissibility filter is entirely in $\Pi_n$, not in $\Omega_n^{(c)}$.
      \end{enumerate}

    \paragraph{Level 4 — The admissibility filter $\Pi_n$.}
      The projection $\Pi_n$ acts on $\Omega_n^{(c)}$ via the BI admissibility
      constraint $A_n \leq c_\chi/\sqrt{\lambda_n}$ and the Gram--Schmidt
      orthogonalisation~\cite{Beau2026a16,Beau2026a26}.
      It selects, from all directions in $\Omega_n^{(c)}$, those compatible with
      the admissibility constraints.
      $\Pi_n$ encodes the laws revealed by the O-series — Born--Infeld bound,
      parity involution, projection locking — but does not define $\Omega_n^{(c)}$.

    \paragraph{Level 5 — The observable outcome $O_n$.}
      The observable state $O_n$ is the image of $\Pi_n$ applied to the admissible
      part of $\Omega_n^{(c)}$.
      It is a discrete, shell-indexed element of the observable
      space~\cite{Beau2026a26}:
      \begin{equation}
        O \;=\; \bigsqcup_{n \in \mathbb{N}_q} O_n.
      \end{equation}

      \begin{remark}[Summary: the five-level hierarchy]
  The five levels constitute a single construction with no external inputs:
  \begin{equation}
    G_q
    \;\xrightarrow{\rho_c}\;
    \mathbb{C}^q
    \;\xrightarrow{\mathrm{BFS}}\;
    \Omega_n^{(c)}
    \;\xrightarrow{\Pi_n}\;
    O_n,
    \qquad
    V_{n-1}^{(c)} \subseteq \Omega_n^{(c)} \subseteq \mathbb{C}^q.
  \end{equation}
  $G_q$ is the relational structure.
  $\mathbb{C}^q$ is its representation carrier.
  $\Omega_n^{(c)}$ is the emergent space of structurally compatible continuations
  (not yet filtered by admissibility),
  constrained from below by the already realized subspace $V_{n-1}^{(c)}$.
  $\Pi_n$ is the admissibility filter.
  $O_n$ is the resolved observable state.
  No level presupposes the next: each is induced by the preceding structure
  and the admissibility constraints.
  This is the concrete answer to the question of where $\Omega_n$ lives:
  in $\mathbb{C}^q$, itself induced by $G_q$ — not in an external configuration
  space.
      \end{remark}

      \begin{remark}[The conjugate pair and $\Omega_n$ without preferred direction]
  The parity involution $c \leftrightarrow q-c$ (O18) acts on $\Omega_n^{(c)}$
  via complex conjugation~\cite{Beau2026a21}:
  $\Omega_n^{(q-c)} = \overline{\Omega_n^{(c)}}$.
  The minimal admissible fibre is therefore the pair
  $\{V_n^{(c)}, V_n^{(q-c)}\} = \{V_n^{(c)}, \overline{V_n^{(c)}}\}$,
  with no preferred orientation — consistent with the admissibility principle
  of A3.
      \end{remark}

  \subsection{From $\mathbb{C}^q$ to continuous spacetime: the large-$q$ limit}
  \label{subsec:continuum-limit}

  The five-level hierarchy of Section~\ref{subsec:five-levels} is formulated
  for finite prime $q$.
  Physical spacetime is continuous.
  This subsection identifies the structural bridge between the two, without
  claiming to complete the derivation.

  \paragraph{The discrete structure at finite $q$.}
    At finite $q$, the observable space $O = \bigsqcup_{n \in \mathbb{N}_q} O_n$
    is discrete and finite.
    The ambient representation space $\mathbb{C}^q$ has dimension $q$.
    Physical quantities are functions on $\mathbb{Z}/q\mathbb{Z}$.

  \paragraph{The large-$q$ limit.}
    As $q \to \infty$ with the prime structure retained, three things happen
    simultaneously:
    \begin{enumerate}[(i)]
\item \emph{The representation space $\mathbb{C}^q$ approaches $L^2(\mathbb{R})$}.
Functions on $\mathbb{Z}/q\mathbb{Z}$ with fixed Fourier support become
functions on $\mathbb{R}$ in the Pontryagin dual limit.
The Weil representation $\rho_c$ of $G_q$ approaches the metaplectic
representation of the real Heisenberg group $\mathrm{Heis}_3(\mathbb{R})$.
\item \emph{The BFS shell structure approaches a continuous foliation}.
The polynomial ball growth $|B_n| \sim c\, n^4$ (Bass--Guivarc'h,
homogeneous dimension $D = 4$) induces, in the $q \to \infty$ limit,
a continuous layering of the configuration space compatible with
a four-dimensional effective geometry~\cite{Cosmochrony}.
\item \emph{The discrete admissibility filter $\Pi_n$ approaches a
differential constraint}.
The Born--Infeld admissibility bound $A_n \leq c_\chi/\sqrt{\lambda_n}$
becomes, in the continuum limit, the Born--Infeld Lagrangian constraint
on effective field configurations~\cite{Beau2026c}.
    \end{enumerate}

    The continuum spacetime observable is therefore not a separate postulate:
    it is the large-$q$ limit of the discrete observable space $O$, with the
    four-dimensional structure inherited from the homogeneous dimension of $G_q$.

    \begin{remark}[Status of the continuum limit]
      \label{rem:continuum-limit}
      The convergence $\mathbb{C}^q \to L^2(\mathbb{R})$ and the emergence of
      continuous spacetime geometry from the BFS shell structure are established
      qualitatively in~\cite{Cosmochrony}.
      The companion paper~\cite{Beau2026q5a} initiates the rigorous treatment of this
      limit, formulating it in the category of Hilbert spaces and operator algebras
      and decomposing it into three distinct convergences: a Hilbert inductive limit
      (T1), a representational convergence of Weil generators to metaplectic
      generators (T2), and a Mosco convergence of admissibility forms to a
      second-order operator $L_\Pi$ on $L^2(\mathbb{R})$ (T3).
      The continuum limit is reduced to a single spectral tightness condition
      (Conjecture~C of~\cite{Beau2026q5a}), which is confirmed quantitatively across all
      tested primes: the spectral tail mass of admissible test vectors satisfies
      $E_q(q/3)/\|f_q\|^2 < 2.1 \times 10^{-5}$ uniformly, with values decreasing
      toward machine precision as $q$ grows.
      The embedding hypothesis~[H1] is proved in the low-frequency regime relevant
      to admissible sequences; the Weil-block compatibility condition and the identification of the effective
      metric from the symbol of $L_\Pi$ are established in the companion paper
      Q5b~\cite{Beau2026q5b}.
      The present remark identifies the mechanism; \cite{Beau2026q5a} establishes the
      operator-theoretic framework and the numerical evidence.
    \end{remark}

    \begin{remark}[Why four dimensions]
  The homogeneous dimension $D = 4$ of the Heisenberg group $G_q$
  (Bass--Guivarc'h theorem: $|B_n| \sim n^4$) is not chosen to match
  the observed four dimensions of spacetime.
  It is a structural consequence of the nilpotency class and rank of
  $\mathrm{Heis}_3$: the group has three generators with one non-trivial
  commutator, giving $D = 2 \cdot \mathrm{rank} + \mathrm{class} = 4$.
  The coincidence with the observed dimension of spacetime is therefore
  a structural prediction of the framework, not a parameter fit.
    \end{remark}

    \clearpage
  \section{Structural Consequences of Non-Injectivity}
  \label{sec:ni-consequences}

  Non-injectivity (A2) and the existence of distinct successor states together
  imply three structural facts whose sources are distinct.

  \subsection{Irreversibility}

    \begin{proposition}[Irreversibility from non-injectivity]
      \label{prop:irreversibility}
      Under A2, every admissible transition involves an irreversible loss of
      structural information.
    \end{proposition}

    \begin{proof}
      Since multiple directions from $O_{n-1}$ map to the same $O_n$ under
      $\Pi_n$, the information distinguishing these directions is absent from
      $O_n$.
      No subsequent transition $\Pi_{n'}$ for $n' > n$ can recover this
      information: the map $\Pi_n$ has positive projection entropy
      $S_{\Pi_n} > 0$~\cite{Beau2026l}, and no admissible transition can reduce
      $S_\Pi$ below zero.
      The loss is therefore structural and permanent;
      its monotone information-theoretic quantification is given in~\cite{Beau2026fe}.
    \end{proof}

  \subsection{Novelty and structural fluctuations}
    \label{subsec:ni-novelty}

    The existence of a distinct successor state $O_n \neq O_{n-1}$ defines a
    structural residue:
    \begin{equation}
      \label{eq:delta-n}
      \Delta_n \;:=\; \mathrm{Res}(O_{n-1}, O_n;\, \Pi_n),
    \end{equation}
    the measure of how much $O_n$ is irreducible to a continuation of $O_{n-1}$
    under $\Pi_n$.
    That $\Delta_n$ is non-trivial is not an additional assumption — it is the
    condition under which we index $O_n$ as a distinct state.
    These residues are what physical descriptions call fluctuations or quantum noise.

  \subsection{Temporal order}

    \begin{theorem}[Temporal order from A2]
      \label{thm:temporal-order}
      Under A2, the chain of observable states $\{O_n\}_{n \geq 0}$ carries a
      canonical oriented, non-trivial partial order.
    \end{theorem}

    \begin{proof}
      Irreversibility (Proposition~\ref{prop:irreversibility}) establishes an
      asymmetric relation: $O_{n-1} \to O_n$ is not reversible, so $O_{n-1}$
      and $O_n$ are not equivalent under the transition structure.
      Novelty ($\Delta_n \neq 0$) ensures $O_n \neq O_{n-1}$, so the order
      is non-trivial.
      The composition of transitions defines transitivity.
      Together: asymmetry, irreflexivity, and transitivity give a strict partial
      order compatible with the chain structure.
    \end{proof}

    \begin{remark}[Time as derived structure]
  The arrow of time is not postulated.
  It is the oriented order generated by the chain of non-injective transitions
  under A2.
  The index $n$ acquires its temporal meaning from this order — not the reverse.
  The quantitative structure of this projective temporal ordering — cumulative
  projected capacity and its numerical validation from the spectral admissibility
  cascade — is developed in~\cite{Beau2026pto}.
    \end{remark}

    \begin{remark}[Admissible transitions as a state machine]
      \label{rem:state-machine}
      The admissible neighbourhood $\mathcal{A}(O_{n-1})$
      (Remark~\ref{rem:admissibility-relational}) gives the
      chain~\eqref{eq:recursive-chain} the structure of an oriented state machine:
      nodes are admissible projections, edges are admissible transitions, and a
      physical history is a directed path in this graph.
      When $|\mathcal{A}(O_{n-1})| = 1$, the successor is uniquely determined by the
      admissibility structure alone — no external law is required.
      When $|\mathcal{A}(O_{n-1})| > 1$, branching occurs; the relative admissibility
      weights of the branches are the projective analogues of quantum probabilities
      (Section~\ref{sec:admissibility}).
      In both cases, no temporal parameter external to the transition structure is
      needed: the index $n$ acquires its temporal meaning entirely from the
      orientation of the admissibility graph.
      \emph{Dynamics is the oriented topology of the admissibility graph, not a law
      imposed upon a pre-existing time.}
    \end{remark}

    \begin{remark}[Observable reality as projective position]
      \label{rem:projective-position}
      The admissible neighbourhood $\mathcal{A}(O_{n-1})$ is not a fixed space of
      possibilities.
      It shifts at every step, because $O_n$ conditions $\mathcal{A}(O_n)$, which differs
      generically from $\mathcal{A}(O_{n-1})$: the projection following step $n$ is never
      performed from an invariant reference frame.
      Observable reality is therefore not a static projection of a fixed substrate onto a
      fixed observable space.
      It is the \emph{current projective position within the space of admissible reprojections}:
      the trajectory $O_0 \leadsto O_1 \leadsto \cdots \leadsto O_n$ is constitutive of
      what is real at step $n$, not contingent upon it.
      The cumulative index $I(n)$ measures not elapsed time but accumulated projective
      divergence from the initial position — how far the sequence has moved through the
      space of admissible neighbourhoods.
      In particular, the projection does not reveal a pre-existing reality: it redefines the
      context from which the next reality may appear.
    \end{remark}

  \subsection{Projective incompleteness}
    \label{subsec:projective-incompleteness}

    Theorem~\ref{thm:heisenberg} establishes that the admissibility structure forces
    $[X, \sigma(X)] = Z \neq 0$.
    This non-commutativity has an immediate structural consequence that is logically
    prior to — and independent of — the wave and phase structure of
    Section~\ref{sec:admissibility}.

    \begin{corollary}[Projective incompleteness]
      \label{cor:projective-incompleteness}
      Under A1--A3 and Born--Infeld parity, no admissible projection $\Pi_n$ resolves
      $\Omega_n^{(c)}$ completely.
      Every admissible projection has a non-trivial kernel and leaves a strictly
      positive informational residue:
      \begin{equation}
        \label{eq:projective-incompleteness}
        \ker \Pi_n \;\neq\; \{0\},
        \qquad
        S_{\Pi_n} \;>\; 0,
      \end{equation}
      for every admissible $\Pi_n$.
    \end{corollary}

    \begin{proof}
      By Theorem~\ref{thm:heisenberg}, $[X, \sigma(X)] = Z \neq 0$.
      A projection $\Pi_n$ with $\ker \Pi_n = \{0\}$ would simultaneously resolve both
      $X$ and $\sigma(X)$ without residual ambiguity, requiring their simultaneous
      diagonalisation and hence $[X, \sigma(X)] = 0$.
      This contradicts $Z \neq 0$.
      Therefore $\ker \Pi_n \neq \{0\}$, and the projection entropy $S_{\Pi_n} > 0$
      follows from Proposition~\ref{prop:irreversibility}.
    \end{proof}

    \begin{remark}[Structural fluctuations as necessary, not contingent]
      \label{rem:fluctuations-necessary}
      Corollary~\ref{cor:projective-incompleteness} changes the epistemic status of the
      structural fluctuations introduced in Section~\ref{subsec:ni-novelty}.
      They are not merely a consequence of abstract non-injectivity (A2 alone), but a
      \emph{necessary consequence of the non-commutativity forced by A1--A3}: any
      projection that left no residue would require $[X, \sigma(X)] = 0$, contradicting
      Theorem~\ref{thm:heisenberg}.
      The quantum vacuum is therefore not a state from which all fluctuations have been
      removed; it is the state from which complete removal is structurally forbidden.
    \end{remark}

    \begin{remark}[Projective incompleteness and the first transition]
      \label{rem:first-transition}
      Corollary~\ref{cor:projective-incompleteness} also resolves a potential question
      about the origin of the first admissible transition: what initiates $\Pi_1$ from
      $\Pi_0$?
      A perfectly self-closed initial state — one with $\ker \Pi_0 = \{0\}$ — is
      structurally forbidden.
      The initial projection necessarily leaves a non-zero residue, which constitutes
      the first difference and guarantees that $\mathcal{A}(O_0)$ is non-empty.
      No external mechanism is required: the first transition is a structural necessity,
      not a contingent event.
      This is the projective analogue of the impossibility of a perfectly symmetric
      initial state: the admissibility constraints that select the Heisenberg structure
      also ensure that a featureless origin is inadmissible.
    \end{remark}

  \section{Admissibility, Wave, and Phase Coherence}
  \label{sec:admissibility}

  \subsection{The admissibility principle}

    Axiom A3 introduces a physical constraint on transitions:

    \begin{quote}
      A transition is admissible if and only if it preserves the full multiplicity
      of open admissible directions without selecting among them prematurely.
    \end{quote}

    Premature selection means making a determinative choice that nothing in the
    current state $O_{n-1}$ justifies.
    Admissibility forbids it: the transition must remain structurally neutral among
    all open directions until projection locking (A4) forces resolution.

  \subsection{The proto-state}
    \label{subsec:proto-state}

    \begin{definition}[Proto-state]
      \label{def:proto-state}
      The \emph{proto-state} $\Psi_n$ associated with a transition from $O_{n-1}$
      to $O_n$ is the unresolved configuration shared between $O_{n-1}$ and $O_n$:
      the carrier of the transition, belonging fully to neither endpoint.
    \end{definition}

    The proto-state is physically real — it is the configuration of the system
    during the transition, before projection locking.
    It is not an epistemic representation of uncertainty about a definite state:
    it is the actual structural state of the system between two realized
    observable states.

    \begin{proposition}[Structural indeterminacy of the proto-state]
      \label{prop:indeterminacy}
      Under A1 and A3, the proto-state $\Psi_n$ is genuinely indeterminate among
      the admissible directions in $F_n$.
    \end{proposition}

    \begin{proof}
      By A1, $F_n$ is non-empty.
      By A3, the transition does not select prematurely among the elements of $F_n$.
      Therefore $\Psi_n$ is not identified with any particular element of $F_n$
      before projection locking: it is open across all of $F_n$ simultaneously.
      This is structural indeterminacy, not epistemic ignorance.
    \end{proof}

  \subsection{Wave, phase, and coherence}

    The proto-state, being structurally open across all elements of $F_n$
    simultaneously, propagates through all admissible directions at once.
    This is the physical origin of wave-like behaviour: not a metaphor, but a
    direct consequence of admissibility.

    A configuration propagating through multiple directions simultaneously carries
    relative phase structure between those directions.
    Phase is therefore not postulated — it is a consequence of structural
    indeterminacy.

    Admissibility (A3) preserves this phase structure: premature selection would
    collapse the relative phases, but A3 forbids it.
    The chain of consequences is therefore:

    \begin{equation}
      \begin{aligned}
        \text{A3}
        &\;\Longrightarrow\; \text{structural indeterminacy} \\
        &\;\Longrightarrow\; \text{proto-state (wave)} \\
        &\;\Longrightarrow\; \text{phase} \\
        &\;\Longrightarrow\; \text{coherence preserved.}
      \end{aligned}
    \end{equation}

    \begin{remark}[The admissible fibre as precursor of Hilbert structure]
  The fibre $F_n$, equipped with the phase structure forced by A3, is the
  natural precursor of an effective Hilbert sector: it carries complex
  amplitudes and interference structure.
  This is not yet a full Hilbert space — the inner product, completeness,
  and probability measure require additional derivation (see Q1, Q2 in
  Section~\ref{sec:summary}).
  But the amplitude structure and interference follow from A3 alone.
    \end{remark}

    \begin{remark}[Double-slit experiment]
  In the double-slit experiment, the proto-state propagates through both slits
  because the admissibility constraints do not select between them:
  both are open directions in $F_n$.
  The interference pattern is a direct consequence of coherent propagation
  through all admissible directions.
  The impact on the screen is the moment the proto-state reaches the
  Born--Infeld saturation threshold — projection locking (A4) forces it onto
  an available shell, resolving it into a localized state $O_n$.
    \end{remark}

  \subsection{Non-factorisability and the emergence of Heisenberg structure}
    \label{subsec:heisenberg}

    Axiom A3 does not only enforce phase coherence.
    It also constrains the algebraic structure of the admissible fibre $F_n$ itself,
    forcing it to carry a Heisenberg structure rather than an arbitrary unitary
    symmetry.

    \begin{remark}[The structural tension between A2 and A4]
      \label{rem:a2-a4-tension}
      The non-factorisability established in this section can be understood as the minimal
      consistent resolution of a structural tension between A2 and A4.

      A2 asserts that the multiplicity of admissible configurations in~$F_n$ cannot be
      eliminated: distinct configurations lead to the same observable, and this
      indistinguishability is structural, not epistemic.
      A4 asserts that the observable outcome is unique at locking.
      Together, they impose an irresolvable constraint: the multiplicity cannot disappear
      (A2 would be violated), yet it cannot remain directly observable (A4 would be violated).

      A3 --- admissibility as non-premature selection --- is the operative axiom that
      determines how this tension resolves.
      It forbids any factorisation of~$F_n$ that would make the multiplicity distinguishable
      prior to locking, forcing it to organise into a globally coherent, irreducible structure.
      Lemma~\ref{lem:non-factorisation} formalises this constraint;
      Theorem~\ref{thm:heisenberg} identifies the resulting structure as Heisenberg type.
      Quantum structure is therefore the minimal algebraic form compatible with the joint
      constraint of A2, A3, and A4.
    \end{remark}

    \begin{lemma}[Admissible non-factorisability]
      \label{lem:non-factorisation}
      Under A1--A3, any decomposition
      \[
        F_n \;=\; F^{(1)} \oplus F^{(2)}
      \]
      that is stable under the admissible generators of $F_n$ is forbidden prior
      to projection locking.
      In particular, A3 forces the action of admissible generators on $F_n$ to
      be irreducible.
    \end{lemma}

    \begin{proof}
      A3 forbids premature selection among open directions.
      A decomposition of $F_n$ stable under admissible generators separates
      directions by invariants that are internal to the fibre — invariants carried
      by the generators themselves, which are constitutive of the admissibility
      structure by A1.
      Such a decomposition therefore resolves the proto-state partially, by
      assigning each open direction to a definite sector, prior to any
      Born--Infeld saturation event.
      This contradicts A3.
      Irreducibility follows immediately: any proper stable sub-representation
      would yield such a forbidden decomposition.
    \end{proof}

    \begin{theorem}[A1--A3 and BI parity force Heisenberg structure]
      \label{thm:heisenberg}
      Under A1--A3 and Born--Infeld parity, the symmetry group of the admissible
      fibre $F_n$ is isomorphic to $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$ for
      some prime $q$.
      Consequently, $F_n$ is isomorphic to the Weil representation $V_\rho$ on
      $L^2(\mathbb{Z}/q\mathbb{Z})$.
    \end{theorem}

    \begin{proof}[Proof sketch]
      Let $X$ and $\sigma(X)$ be the conjugate generators of $F_n$ under
      Born--Infeld parity.
      If $[X, \sigma(X)] = 0$, then $F_n$ decomposes into simultaneous eigenspaces,
      yielding a stable decomposition forbidden by Lemma~\ref{lem:non-factorisation}.
      Hence $[X, \sigma(X)] \neq 0$.
      Minimality (A1) forces the commutator $Z := [X, \sigma(X)]$ to be central:
      any non-central $Z$ would extend the algebra beyond the conjugate pair,
      contradicting the minimal generative closure forced by A1.
      The group $\langle X, \sigma(X) \rangle$ is then a non-trivial central
      extension of $(\mathbb{Z}/q\mathbb{Z})^2$ by $\langle Z \rangle$, isomorphic
      to $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$ by the classification of such
      extensions.
      Lemma~\ref{lem:non-factorisation} further implies that $|Z|$ is prime: if
      $|Z|$ were composite, the representation of $F_n$ would decompose over the
      characters of $\langle Z \rangle$, yielding a forbidden stable decomposition.
      Finally, Born--Infeld parity forces $c \neq 0$~\cite{Beau2026c}, so the
      central character is non-trivial; the discrete Stone--von Neumann theorem then
      identifies $F_n \simeq V_\rho$ uniquely.
      The detailed proof of each step is given in the companion
      paper~\cite{Beau2026HeisenbergStructure}.
    \end{proof}

    \begin{equation}
      \begin{aligned}
        \rlap{\text{A3}}\phantom{\text{BI parity}}
        &\;\Longrightarrow\; \text{no admissible factorisation}
        \;\Longrightarrow\; [X, \sigma(X)] \neq 0
        \;\Longrightarrow\; \text{central extension,} \\
        \rlap{\text{A1}}\phantom{\text{BI parity}}
        &\;\Longrightarrow\; \text{minimal generative closure}
        \;\Longrightarrow\; \text{Heisenberg algebra,} \\
        \rlap{\text{A3}}\phantom{\text{BI parity}}
        &\;\Longrightarrow\; \text{irreducibility}
        \;\Longrightarrow\; |Z| = q \text{ prime}
        \;\Longrightarrow\; \mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z}), \\
        \text{BI parity}
        &\;\Longrightarrow\; c \neq 0
        \;\Longrightarrow\; F_n \;\simeq\; V_\rho.
      \end{aligned}
    \end{equation}

    \begin{remark}[Uncertainty principle as a corollary]
      \label{rem:uncertainty}
      Theorem~\ref{thm:heisenberg} subsumes the open direction~Q6 of
      Section~\ref{sec:summary}.
      The incompatibility of conjugate observables $X$ and $\sigma(X)$ — i.e.\
      the Heisenberg uncertainty principle — is a direct corollary: it follows
      from $[X, \sigma(X)] \neq 0$ together with the standard operator inequality.
      The uncertainty principle is therefore not a property of the observables,
      but a property of the non-resolved proto-state: it is the algebraic
      signature of the prohibition on premature selection encoded in A3.
    \end{remark}

    \begin{remark}[Canonical quantisation as structural necessity]
      The relation $[X, \sigma(X)] = Z$ defines a minimal discrete symplectic
      structure, of which the Heisenberg algebra is the canonical quantisation.
      Theorem~\ref{thm:heisenberg} therefore establishes that canonical
      quantisation is not a modelling choice: it is the unique algebraic structure
      compatible with the requirement that the proto-state remain genuinely
      non-resolved until projection locking.
    \end{remark}



    A quantum superposition is an unresolved proto-state: a physically real
    configuration whose fibre maintains coherent relations between multiple open
    admissible directions.
    There is no ontological paradox — superposition is the normal condition of
    a system between two transitions.

    A measurement is a transition whose admissibility constraints are strong enough
    to force resolution of the proto-state onto a single admissible direction.
    Free evolution and measurement differ in the strength of their admissibility
    constraints, not in their fundamental mechanism.

    In this sense, quantum mechanics is the effective theory of physically real
    unresolved admissible structures and their resolution.

  \subsection{The wavefunction as effective representation of the proto-state}
    \label{subsec:wavefunction}

    A quantum wave is not a physical substance propagating in space.
    It is the effective representation of a physically real unresolved proto-state.
    It encodes the coherent structure of admissible directions prior to projection
    locking.

    More precisely: the proto-state $\Psi_n$ is real;
    the wavefunction $\psi$ is a coordinate description of the coherent admissible
    structure carried by $\Psi_n$ in the spectral realisation $V_\rho$ of $F_n$.
    When we write $\psi$, we are not describing a substance propagating in space
    — we are encoding the relational phase structure of an unresolved
    configuration in the space of admissible directions.
    Quantum mechanics does not describe realized physical states:
    it describes states in the process of being resolved.

    This resolves the standard interpretational difficulties without additional
    postulate:
    \begin{itemize}
      \item \emph{Superposition} is the normal condition of a proto-state — the
      simultaneous structural openness across all elements of $F_n$.
      \item \emph{Interference} is the direct consequence of coherent propagation
      through multiple admissible directions before resolution.
      \item \emph{Collapse} is not a physical process applied to a pre-existing state
      but the resolution of the proto-state at projection locking — the transition
      from unresolved admissible structure to a realized observable outcome.
    \end{itemize}

  \subsection{Why complex amplitudes}
    \label{subsec:complex-amplitudes}

    Given that the proto-state is a physically real, unresolved, coherent
    configuration, what is the minimal mathematical structure capable of
    representing it?

    Real positive weights are insufficient.
    They describe ignorance about which of several pre-existing alternatives is
    realized — a classical mixture.
    The proto-state is not a mixture: it is a single coherent configuration
    with definite phase relations between admissible directions.

    Real signed weights are also insufficient.
    They allow opposition of sign but do not support a rich enough phase structure
    under composition.
    In particular, they cannot represent the full coherence of a proto-state whose
    directions combine under the SU(2) structure forced by O23.

    The minimal structure that can:
    (i) represent simultaneous coherent openness across multiple directions,
    (ii) encode their relative phase relations,
    (iii) compose correctly under the SU(2) symmetry of the admissible fibre,
    (iv) support interference — the physical consequence of coherence before
    resolution —
    is a complex amplitude.

    Therefore: complex amplitudes are not introduced by convention.
    They are the minimal representation of a coherent unresolved proto-state.
    The imaginary unit encodes the relative phase between admissible directions;
    it enters because the proto-state is not a mixture but a coherent
    superposition.

  \subsection{Why the Born weight is the squared norm}
    \label{subsec:born-weight}

    Upon projection locking (A4), the unresolved proto-state is resolved onto
    a single admissible direction.
    The quantity associated with each direction at resolution — the
    \emph{intensity of realisation} — must satisfy four structural requirements:

    \begin{enumerate}[(i)]
\item \emph{Positivity}: a weight on admissible directions cannot be negative.
\item \emph{Global phase invariance}: multiplying $\psi$ by a global phase
$e^{i\theta}$ does not change the physical proto-state, so the weight must
be invariant under this transformation.
\item \emph{Normalisability}: the total weight over all admissible directions
must be finite and summable to one.
\item \emph{Coherence preservation before resolution}: the weight must not
erase phase relations prematurely — interference must remain possible up to
the moment of resolution.
    \end{enumerate}

    The squared norm $|\psi|^2$ is the unique minimal quantity satisfying all
    four requirements.
    The norm $|\psi|$ fails (ii) under phase shifts and (iv) — it introduces
    real-valued interference which is not the correct structure.
    Any higher power $|\psi|^p$ for $p \neq 2$ fails (iii) under composition
    of superpositions.
    Any non-analytic function of $\psi$ fails (iv).

    The Born weight is therefore not a postulate.
    It is the positive intensity associated with the resolution of a coherent
    admissible structure — the unique minimal such intensity compatible with
    the four structural requirements.

    \begin{remark}[Relation to the Born rule derivation in the companion paper]
  The argument above establishes the structural necessity of the squared norm
  as the resolution intensity of the proto-state.
  The companion paper~\cite{Beau2026q1} makes this precise as a theorem in the SU(2)
  parity sector: the Born rule $P(a=+1|\hat{a}) = |\langle +\hat{a}|\psi\rangle|^2$
  is the unique probability assignment compatible with the coherent fibre
  structure, the SU(2) observable algebra, and the derived singlet correlator.
  The two derivations are complementary: the present argument gives the
  structural motivation; the companion gives the uniqueness proof.
    \end{remark}

    \begin{remark}[The wavefunction lives in the admissible fibre, not in space]
  A common source of confusion is to picture the wavefunction as an object
  propagating in physical space.
  In the present framework, $\psi$ encodes the structure of $F_n$ — the
  admissible directions from $O_{n-1}$.
  Its concrete realisation is the spectral sector $V_\rho$ induced by the
  admissibility constraints on the Heisenberg Cayley graph (O-series).
  $F_n$ and $V_\rho$ are not pre-existing spaces: they are the spaces
  defined by the admissibility constraints.
  The wavefunction does not propagate in space; it encodes the relational
  phase structure of an unresolved admissible configuration.
    \end{remark}


    \label{sec:locking}

    Axiom A4 establishes that resolution is discrete.
    This discreteness is not postulated: it follows from the interaction of two
    structures already present in the framework.

  \subsection{Two types of discreteness}

    The Born--Infeld saturation condition~\cite{Beau2026c} operates on a
    continuous amplitude function.
    Its saturation locus $\mathcal{L}_{\mathrm{BI}} \subset \mathbb{R}$ is
    generically real-valued — its elements carry no information about shell
    structure.

    The observable space decomposes discretely:
    \begin{equation}
      O \;=\; \bigsqcup_{n \in \mathbb{N}_q} O_n,
    \end{equation}
    indexed by BFS shells of the Heisenberg Cayley graph~\cite{Beau2026a26}.
    No component $O_{n'}$ exists for non-admissible depths.

    The proto-state is resolved at the first element of
    $\mathcal{L}_{\mathrm{BI}} \cap \mathbb{N}_q$ — the first depth where
    Born--Infeld saturation coincides with an available shell.
    This is the Projection Locking theorem~\cite{Beau2026a26}.

    \begin{theorem}[Discreteness from projection locking, after O22]
      \label{thm:discrete-locking}
      Under A3 and A4, quantum transitions are discrete.
      The discreteness is a consequence of the interaction of continuous
      Born--Infeld saturation with the discrete admissibility filter of $\Pi_n$,
      not a postulate.
    \end{theorem}

  \subsection{Three stable directions}

    The number of independent stable admissible directions in the fibre is
    exactly three, derived algebraically from the admissibility constraints
    alone~\cite{Beau2026a27}:
    \begin{equation}
      \begin{aligned}
        \text{BI parity}
        &\;\Rightarrow\; \text{fibre involution } c \leftrightarrow q-c \\
        &\;\Rightarrow\; \text{non-abelian admissible support} \\
        &\;\Rightarrow\; \mathbb{H}\text{-minimality} \\
        &\;\Rightarrow\; \dim_{\mathbb{R}} \mathrm{Im}\,\mathbb{H} = 3 \\
        &\;\Rightarrow\; \Sigma_c(n_3) = 3.
      \end{aligned}
    \end{equation}
    The emergence of exactly three stable directions — and with it, three
    particle generations — is algebraic in origin, not geometric or
    phenomenological.

    \clearpage
  \section{Entanglement and Bell Non-Applicability}
  \label{sec:entanglement}

  \subsection{Entanglement as a shared infra-projectable cell}

    Two observable systems $A$ and $B$ are entangled when their joint description occupies a single
    non-decomposable cell of the projection $\Pi_n$: a fibre that $\Pi_n$ cannot factorise into separate
    projected descriptions of $A$ and $B$.
    The correlation between them is exactly the part of the joint fibre that $\Pi_n$ cannot split.

    This is a synchronic, structural property of the non-injective projection, not a residue of a past
    interaction.
    The binding is the non-factorisability of the fibre forced by A1 and A3 (Section~\ref{sec:admissibility});
    it holds wherever a shared non-decomposable cell is present, independently of how that cell arose.
    For the conjugate Weil pair $\{c,q-c\}$ that organises the spectral cascade, the binding is explicit and
    permanent: the two characters are related by the Born--Infeld parity involution
    $V_{q-c}\simeq\overline{V_c}$, which holds at every depth $n$.
    The pair is a single closed object throughout the cascade, not two subsystems that were joined at an
    earlier transition and have since separated.

    This picture is now quantitatively derived rather than asserted.
    For the conjugate pair, the minimal admissible projection is spectrally atomic: after removal of the zero
    mode it is the two-dimensional Born--Infeld parity pair, identified with the spin-$\tfrac12$
    representation, and its reduced state is the singlet with entanglement entropy
    $\log 2$~\cite{Beau2026ent2,Beau2026q3}.
    This is the entropy of the state behind the singlet correlator and the Tsirelson bound~\cite{Beau2026q1}.
    A finer pipeline-level observable, the full residual fibre-level rank $r_{\mathrm{pair}}(n)$, gives
    $S_{\mathrm{ent}}(n)=\log r_{\mathrm{pair}}(n)$, proved for matched single-character blocks and conditional
    otherwise on a residual indiscernibility hypothesis~\cite{Beau2026entnote,Beau2026ent1}.
    The two observables must not be conflated; they coincide only when $r_{\mathrm{pair}}(n)=2$.

    The invariant carried by the shared cell is infra-projectable: it has no position in the projected
    temporal ordering, which is itself the cumulative projected capacity.
    It is a single unsplit object below the resolution of $\Pi_n$ (Corollary~\ref{cor:projective-incompleteness}),
    not a pair of separately ordered events.
    In this sense entanglement is the not-yet-time-ordered sector of the cascade, rather than a record of a past
    joining.
    Correlations of this kind carry no signal and admit no influence between separated projected descriptions;
    locality at the projected level is preserved.

  \subsection{Why Bell's factorisation assumptions do not apply}

    Bell's theorem~\cite{Bell1964} constrains models in which measurement outcomes are determined by a variable
    $\lambda$ satisfying:
    (i) $\lambda$ is local (assignable separately to $A$ and $B$),
    (ii) $\lambda$ is a definite value present at measurement,
    (iii) $\lambda$ is independent (statistically independent of measurement settings).

    The fibre invariant in the present framework satisfies none of these conditions, and the decisive failure
    is (ii).
    A variable fixed in the common past history would itself be a local hidden variable of exactly the kind
    Bell's theorem excludes; locating the shared fibre in the past would therefore not evade the theorem.
    The escape is instead that the fibre invariant is infra-projectable: it never acquires a definite projected
    value, at the present transition or any other, because it lives below the resolution of $\Pi_n$
    (Corollary~\ref{cor:projective-incompleteness}).
    There is no $\lambda$ with a definite value to be assigned.

    The invariant is also not local --- it belongs to the shared non-factorisable cell, not to a separable
    region assignable to $A$ or $B$ --- and not independent, since $A$ and $B$ occupy the same cell, which
    cannot be factorised.
    The Bell factorisation assumption $P(a,b|\lambda) = P(a|\lambda) \cdot P(b|\lambda)$ fails because it
    presupposes a definite $\lambda$ assignable separately to $A$ and $B$, and the shared infra-projectable
    cell provides neither.
    Bell-type inequalities are therefore not violated by a hidden mechanism in this framework --- their
    preconditions do not hold.
    This is established formally in the companion paper~\cite{Beau2026b}.

    \clearpage
  \section{Summary: What Is Derived from What}
  \label{sec:summary}

  Table~\ref{tab:derived} summarises the derivation chain.

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{p{5.2cm}p{5.8cm}p{3.2cm}}
      \toprule
      \textbf{Concept} & \textbf{Source} & \textbf{Status} \\
      \midrule
      \multicolumn{3}{l}{\textit{Direct axiomatic consequences (A1--A4)}} \\
      \midrule
      Irreversibility &
      Non-injectivity (A2) &
      Proved (Prop.~\ref{prop:irreversibility}) \\
      Temporal order &
      Irreversibility + novelty (A2) &
      Proved (Thm.~\ref{thm:temporal-order}) \\
      Proto-state (wave) &
      Unresolved multiplicity (A1, A3) &
      Derived (Prop.~\ref{prop:indeterminacy}) \\
      Phase coherence &
      Admissibility preserves open directions (A3) &
      Derived \\
      Discrete transitions &
      Projection locking (A4, O22) &
      Proved (Thm.~\ref{thm:discrete-locking}) \\
      Three stable directions &
      $\dim_\mathbb{R} \mathrm{Im}\,\mathbb{H} = 3$ (O23) &
      Proved \\
      Entanglement &
      Shared non-factorisable infra-projectable cell~\cite{Beau2026entnote} &
      Derived \\
      Entanglement entropy &
      Admissible pair $\to\log 2$; residual rank $\to\log r_{\mathrm{pair}}$~\cite{Beau2026ent2,Beau2026ent1} &
      Proved / cond. \\
      Bell non-applicability &
      Non-local, infra-projectable fibre &
      Proved~\cite{Beau2026b} \\
      Born rule (SU(2) sector) &
      Correlator + coherence~\cite{Beau2026q1} &
      Proved \\
      $F_n \simeq V_\rho$ (Heisenberg structure) &
      Non-factorisability (A1, A3) + BI parity &
      Proved (Thm.~\ref{thm:heisenberg}) \\
      \midrule
      \multicolumn{3}{l}{\textit{Key results of the companion programme
        (see~\cite{Beau2026a} for the full derivation chain)}} \\
      \midrule
      SU(2) as admissibility fixed point &
      Spectral convergence (Q2)~\cite{Beau2026q2} &
      Proved \\
      Effective spacetime geometry, $D_{\mathrm{hom}}=4$;
      full co-metric $g^{\mu\nu}=\mathrm{diag}(-2,2,2,2)$ &
      BFS stratification (Q5b, Q6b, Q8, Q10, Q11)~\cite{Beau2026q5b,Beau2026q6b,BeauQ8,Beau2026q10,Beau2026q11} &
      Proved \\
      Einstein equation $G_{\mu\nu} = 8\pi G\,T_{\mu\nu}$ &
      Spectral entropy variation~\cite{Beau2026h} &
      Proved (structural) \\
      SU(2) gauge structure &
      Quaternionic rigidity + Q6a~\cite{Beau2026a31,Beau2026q6a} &
      Proved \\
      Yang--Mills equations $D_\mu F^{a\mu\nu} = 0$ &
      Spectral entropy $a_4$ variation (Q12)~\cite{Beau2026q12} &
      Proved (structural) \\
      Gauge--matter coupling $G_{\mu\nu} = 8\pi G_N T^{YM}_{\mu\nu}$ &
      Joint variation $S_\Pi[g,A]$ (Q13)~\cite{Beau2026q13} &
      Proved (structural) \\
      Hierarchy $G_N g^2_{YM} \ll 1$ without fine-tuning &
      Seeley--DeWitt stratification (Q13)~\cite{Beau2026q13} &
      Structural \\
      Tower of non-injectivity &
      ENI Corollary~6~\cite{Beau2026l} &
      Proved \\
      $G_{\mathrm{SM}} =
      \mathrm{SU}(3)\times\mathrm{SU}(2)\times\mathrm{U}(1)$ &
      Colour co-admissibility (O31, O32)~\cite{Beau2026a35,Beau2026a36};
      single-frequency BI fingerprint &
      Proved for the standard graph
      (O31~v1.5, Prop.~4.23~\cite{Beau2026a35}) \\
      \midrule
      \multicolumn{3}{l}{\textit{Open}} \\
      \midrule
      Full Hilbert structure (continuous spectrum) &
      Beyond discrete fibre; large-$q$ limit &
      \textbf{Open} \\
      Born rule (general observables) &
      Beyond SU(2)/$V_\rho$ sector &
      \textbf{Open} \\
      Colour charges, quark sector &
      Individual sectors $c_i$ (O31, O32) &
      \textbf{Open} \\
      \bottomrule
    \end{tabular}
    \caption{Derivation chain from the four axioms.
    ``Derived'': structural consequence of the axioms.
    ``Proved'': explicit proof in this paper or a cited companion.
    ``Proved (structural)'': established via structural argument
    internally consistent with the framework; full formalisation deferred.
    ``Conditional'': proved subject to a named hypothesis.}
    \label{tab:derived}
  \end{table}

  \subsection*{Partial spectral stratification: an organising pattern from the companion programme}
    \label{ssec:spectral-stratification}

    The four axioms A1--A4 are the axiomatic primitive of the programme.
    They do not by themselves determine the interaction content of the emergent theory;
    that content emerges from the spectral structure of the admissible operator,
    developed in the Q-series companion papers.
    The companion papers~\cite{Beau2026h,Beau2026q12,Beau2026q13} establish
    the following partial pattern, which organises the physical interpretation
    of the Seeley--DeWitt expansion of $S_\Pi[g,A]$:

    \begin{principle}[Spectral stratification of effective interactions]
      \label{prin:stratification}
      The physical role of an emergent interaction is determined not only by the
      symmetry it carries, but by the spectral order at which the admissible operator
      responds to variation.
      The gravitational sector appears at the first non-trivial horizontal response;
      the gauge sector appears at the vertical curvature response;
      and mixed gauge--gravity terms appear at the next coupled order.
    \end{principle}

    The evidence base for Principle~\ref{prin:stratification} is as follows:

    \[
      \begin{array}{c|c|l}
        \text{Spectral order} & \text{Sector} & \text{Geometric origin} \\
        \hline
        a_2 & \text{gravity}
        & \text{horizontal variation of the base metric}~\text{\cite{Beau2026h}} \\
        a_4 & \text{gauge}
        & \text{vertical variation of the fibre connection}~\text{\cite{Beau2026q12}} \\
        a_6 & \text{mixed coupling}
        & \text{joint base--fibre curvature terms}~\text{\cite{Beau2026q13}} \\
        a_{\mathrm{spin}} & \text{fermionic structure}
        & \text{Dirac-type admissible operator (open)} \\
      \end{array}
    \]

    \begin{remark}
      \label{rem:stratification-status}
      Principle~\ref{prin:stratification} is \emph{not} derived from A1--A4 alone.
      Its correct epistemic status is: established by the companion programme,
      consistent with A1--A4, and subject to the conditionalities of each
      companion paper.
      In particular, the identification $G_\Pi = \mathrm{SU}(3)\times\mathrm{SU}(2)\times\mathrm{U}(1)$
      is established for the standard graph via~$[\mathrm{H\text{-}color}]$, proved analytically
      in O31~\cite{Beau2026a35} (Proposition~4.23) via the single-frequency structure of the
      fingerprint: each O12 fingerprint vector carries frequency
      $B = c_1 b_1 + c_2 b_2 + c_3 b_3$, and the orbit dilation $B\mapsto\omega B$ is a
      bijection of $\mathbb{Z}/q\mathbb{Z}$.
      The test of this prediction is whether admissibility forces a Clifford module
      and a $\mathbb{Z}_2$-grading on the admissible operator.
    \end{remark}

  \subsection*{Open directions}

    \paragraph{Q2 — Born rule beyond SU(2).}
      The Born rule $P(a=+1|\hat{a}) = |\langle +\hat{a}|\psi\rangle|^2$ is derived
      in~\cite{Beau2026q1} for the SU(2) parity sector.
      A first step beyond this sector is taken in~\cite{Beau2026q2}, which derives the
      Born rule, singlet correlator, and Tsirelson bound for the co-admissible
      spin-$\tfrac{3}{2}$ sector on the binary icosahedral group $2I$, and
      identifies SU(2) as the unique stable sector under admissibility refinement
      in the large-$p$ limit.
      The full extension to arbitrary observables, multipartite systems, and
      continuous spectra requires extending the fibre structure beyond the
      two-element parity case and remains the primary open direction.

    \paragraph{O31/O32 --- Colour sector and recursive non-injectivity.}
      The SU(2) derivation proceeds via conjugate pairs $\{c, q-c\}$.
      For primes $q \equiv 1 \pmod{3}$, the Weil representation of
      $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$ admits colour triplets
      $\{c_1, c_2, c_3\}$ satisfying $c_1 + c_2 + c_3 \equiv 0 \pmod{q}$
      and pairwise non-conjugacy.
      Under $[\mathrm{H\text{-}color}]$ (triplet co-admissibility,
      proved analytically in O31~\cite{Beau2026a35} via the single-frequency structure of the
      Born--Infeld fingerprint),
      O31~\cite{Beau2026a35} derives $\mathrm{SU}(3)$ as the admissibility
      fixed point of the colour fibre, completing the identification
      $G_{\mathrm{SM}} = \mathrm{SU}(3) \times \mathrm{SU}(2) \times \mathrm{U}(1)$.

      Each element $c_i$ of a valid triplet already carries spin-$\tfrac{1}{2}$
      (from its conjugate pair $\{c_i, q-c_i\}$, O29~\cite{Beau2026a33})
      and one colour charge (from the triplet structure).
      This is the minimal quantum-number content of a quark.
      The identification of quarks as stable projected configurations in
      $\mathrm{Im}\,\Pi$ carrying these quantum numbers, and the derivation of
      colour confinement as a structural consequence of the projection
      $\Pi_{\mathrm{colour}} : \{c_i\} \to \{\text{colour-neutral objects}\}$
      being non-injective (ENI~\cite{Beau2026l}, Corollary~6), are the primary
      open directions of the colour sector.

      More broadly, ENI Corollary~6 establishes that the four axioms force
      non-injectivity not only at the level of the fundamental projection
      $\Pi_1$, but at every stratum of a tower of genuinely distinct emergent
      levels.
      The programme's current architecture instantiates two strata: the
      pair level (leptons, $\Pi_1$) and the triplet level (colour sector,
      $\Pi_2$).
      The structural derivation of the full particle hierarchy from this
      recursive tower is the central open direction of the programme.

    \paragraph{Spectral quasi-fixed points and stable structures.}
      The admissible neighbourhood $\mathcal{A}(O_n)$ varies generically with $n$
      (Remark~\ref{rem:projective-position}).
      However, configurations $O^*$ for which $\mathcal{A}(O^*)$ is approximately
      self-similar --- in the sense that projection from $O^*$ tends to return to a
      spectrally close neighbourhood --- define \emph{spectral quasi-fixed points} of
      the admissible flow.
      These are not static states, which would contradict the irreversibility established in
      Proposition~\ref{prop:irreversibility}, but projective orbits that traverse the same
      spectral region repeatedly.
      Stable physical structures --- particles, gauge symmetries, and coupling constants ---
      are candidates for identification with such quasi-fixed points: their observed
      stability would then be a projective property, arising from the near-invariance of
      their admissible neighbourhood under repeated projection, not from any ontological
      fixity.
      The universality of $a_q(s) \cdot q^2 \approx 2$ across all tested primes
      (O-series numerics) is consistent with the existence of such an attractor in
      the admissible spectral flow, independent of~$q$.
      This numerical observation acquires a structural anchor from the companion
      paper Q10~\cite{Beau2026q10}: the horizontal coefficient $A_H(q)$ of the
      effective spatial co-metric satisfies $A_H(q) \to 2$ as $q \to \infty$, where
      the limiting value is identified as the Casimir eigenvalue
      \begin{equation}
        \label{eq:casimir-2}
        C_{\mathrm{su}(2)}\big|_{j=1} \;=\; j(j+1) \;=\; 2
      \end{equation}
      of the spin-$1$ representation $\mathrm{Sym}^2(V_\rho)$ — the unique
      $\mathrm{su}(2)$-invariant quadratic form on the effective spatial sector
      (Q10, Proposition~5.1, via Schur's lemma).
      The epistemic chain leading to this identification is:
      \begin{equation}
        \label{eq:casimir-chain}
        [X,Y] \neq 0
        \;\Rightarrow\; \text{no perfect resolution}
        \;\Rightarrow\; [\mathrm{U}]
        \;\Rightarrow\; \mathrm{su}(2)\text{-isotropy}
        \;\Rightarrow\; C_{\mathrm{su}(2)}|_{j=1} = 2
        \;\Rightarrow\; A_H(q) \to 2,
      \end{equation}
      where: the first implication is Corollary~\ref{cor:projective-incompleteness};
      the universality hypothesis~$[\mathrm{U}]$ is structurally supported and
      numerically confirmed for $q \leq 211$~\cite{Beau2026a29,Beau2026q10};
      the su(2)-isotropy step is Q10~Lemma~4.1; and the Casimir identification is
      Q10~Proposition~5.1 (unconditional).
      The convergence $A_H(q) \to 2$ is established conditionally on~$[\mathrm{U}]$
      in Q10~Theorem~1.1.
      The attractor value $2$ is therefore not a free parameter but is forced by the
      representation theory of the admissible fibre, providing a
      representation-theoretic interpretation of the quasi-fixed point.
      Formalising the existence and classification of quasi-fixed points of
      $\mathcal{A}$ is an open direction of the programme.

      Theorem~\ref{thm:heisenberg} establishes $F_n \simeq V_\rho$ at finite
      prime~$q$.
      The relationship between this discrete structure and continuous spacetime
      geometry in the large-$q$ limit — i.e.\ how $V_\rho$ on
      $L^2(\mathbb{Z}/q\mathbb{Z})$ approaches $L^2(\mathbb{R})$ and how the
      BFS shell structure produces a four-dimensional effective geometry —
      is addressed in the companion spectral geometry paper~\cite{Beau2026q5a}
      (Section~\ref{subsec:continuum-limit}).
      That paper reduces Q5 to a single spectral tightness condition (Mosco
      tightness of the admissibility forms), confirms this condition
      quantitatively across all tested primes, and proves the embedding in the
      low-frequency regime relevant to admissible sequences.
      The identification of the effective metric from the principal symbol of the
      limit operator $L_\Pi$ and the Lorentzian signature selection are established
      in the companion paper Q5b~\cite{Beau2026q5b}.
      The temporal coefficient $A_\tau = 2$ is subsequently fixed by Q11~\cite{Beau2026q11}
      via temporal Casimir rigidity, closing the effective co-metric to
      $g^{\mu\nu} = \mathrm{diag}(-2,\,2,\,2,\,2)$.

    \paragraph{Q14 --- Fermionic structure and chirality.}
      The spectral stratification of Principle~\ref{prin:stratification} identifies
      the next open level of the programme: the fermionic sector.
      The companion papers Q12 and Q13~\cite{Beau2026q12,Beau2026q13} derive
      bosonic gauge and gravitational dynamics from a Laplace-type admissible operator
      $\mathcal{A}_{g,A}$.
      Fermionic structure requires a Dirac-type operator $\mathcal{D}_{g,A}$
      satisfying $\mathcal{D}^2 = \mathcal{A}_{g,A} + \tfrac{1}{4}R$
      (Lichnerowicz formula), whose $\mathbb{Z}_2$-grading encodes chirality.
      The central open question is whether the admissibility structure forces a
      Clifford module structure on the admissible fibre and thereby selects a
      $\mathbb{Z}_2$-grading: if so, the chiral asymmetry of the Standard Model
      becomes a structural consequence of admissibility; if not, this constitutes
      a structural boundary of the programme.

      \clearpage
  \section{Relation to Established Frameworks}
  \label{sec:frameworks}

  The four axioms A1--A4 constitute a primitive layer from which geometry, dynamics,
  and quantum structure emerge without being presupposed.
  It is natural to ask how this primitive relates to the established formalisms of
  theoretical physics — not in the sense of reduction, but in the sense of structural
  translation: what each framework corresponds to, where it appears in the emergent
  hierarchy, and what it presupposes that the present framework derives.

  \paragraph{Hamilton--Jacobi dynamics.}
    Once the effective metric $g^{\mu\nu}$ has been reconstructed from the principal
    symbol of the admissibility operator (Q5b, Q6b~\cite{Beau2026q5b,Beau2026q6b}),
    the characteristic propagation equation of the effective geometry coincides with the
    Hamilton--Jacobi equation for massless propagation.
    This identification belongs to the already projected geometric regime, downstream of
    the admissibility reconstruction.
    Hamilton--Jacobi theory should therefore be understood as an effective reformulation
    of the projected dynamics — valid once the effective configuration space has been
    constructed by the primitive admissibility relation — not as a primitive layer of the
    framework.
    The detailed identification, including the eikonal equation as principal-symbol
    characteristic, is given in Q6b~\cite{Beau2026q6b}.

  \paragraph{Symplectic geometry and analytical mechanics.}
    The framework of symplectic geometry takes a phase space — a cotangent bundle
    $T^*M$ equipped with a canonical symplectic form — as its fundamental arena.
    In the present framework, this phase space is not a primitive: the effective
    configuration space $M$ emerges from the Carnot--Carath\'eodory geometry of the
    BFS shell stratification (Q5b~\cite{Beau2026q5b}), and the cotangent structure is
    induced by the principal symbol of the admissibility operator.
    Symplectic geometry is therefore the appropriate language for the effective dynamics
    on $T^*M$, valid after the configuration space has been constructed by the
    admissibility constraints; it does not describe the admissibility structure that
    produces it.

  \paragraph{WKB approximation and geometric optics.}
    In the WKB approximation, the quantum wave function takes the form
    $\psi \sim A\,e^{iS/\hbar}$, where the phase $S$ is the Hamilton--Jacobi action.
    In the present framework, the phase of an admissible mode is not an input but an
    output of the fibre structure: it emerges from the coherence conditions imposed by
    the admissibility filter on the Weil representation~\cite{Beau2026q1}.
    The WKB identification therefore holds at the level of the effective description,
    once the continuum limit has been taken (Q5a~\cite{Beau2026q5a}): the collective
    phase of the admissible modes plays the role of $S$, and geometric optics is the
    ray approximation of the admissible propagation structure in this limit.
    The WKB phase is thus a derived quantity, encoding the projective compression of the
    admissible fibre, rather than a fundamental dynamical object.

  \paragraph{Functional renormalisation group.}
    The Wetterich effective average action satisfies an exact renormalisation group
    equation whose infrared fixed point recovers the physical effective action.
    In the present framework, the Mosco limit of the admissibility Dirichlet forms
    (Q5a~\cite{Beau2026q5a}) plays a structurally analogous role: the sequence
    $\{\mathcal{E}_q\}$ flows, as $q\to\infty$, to the scalar operator $\mathcal{L}_\Pi$,
    whose principal symbol defines the effective metric.
    The spectral entropy functional $S_\Pi[g,A]$ of the companion programme is the
    effective action resulting from this limit~\cite{Beau2026h,Beau2026q12,Beau2026q13}.
    The structural difference is that the Wetterich flow renormalises a pre-existing field
    theory, while the present limit produces the field content itself from the admissibility
    constraints.

  \paragraph{Page--Wootters construction and relational time.}
    The Page--Wootters construction derives apparent temporal evolution from conditional
    states of a globally stationary wave function, treating time as a relational label on
    a clock subsystem.
    The projective temporal ordering of the present framework~\cite{Beau2026pto} shares
    this relational spirit — time is not a parameter of the substrate but a monotone
    ordering induced by the admissibility cascade — but differs in one structural respect.
    In the Page--Wootters model, the global state is fully recoverable (the projection is
    effectively injective), whereas here the projection $\Pi$ is structurally non-injective
    (A2, ENI~\cite{Beau2026l}).
    Temporal ordering in the present framework is therefore irreversible at the structural
    level: it reflects the irreversible accumulation of projected spectral capacity, not
    the choice of a clock observable.

  \paragraph{Holographic renormalisation and radial Hamilton--Jacobi.}
    In holographic renormalisation, the on-shell gravitational action satisfies a radial
    Hamilton--Jacobi equation whose flow encodes the renormalisation group of the boundary
    field theory.
    The Schwarzschild geometry derived in Q6b~\cite{Beau2026q6b} from flux conservation
    of the admissibility operator has a structural parallel with this picture: both
    frameworks derive a bulk geometry from a propagation constraint, and both identify the
    horizon as a degeneration of the effective description.
    The structural difference is that holographic renormalisation presupposes a fixed
    anti-de Sitter background and a specific bulk-boundary duality, while the present
    framework derives the effective geometry from the admissibility operator without a
    background or a boundary field theory.

    \clearpage
  \section*{Acknowledgements}

  The author acknowledges the use of large language models as a supportive tool
  for refining language, structure, and internal consistency during the development
  of this manuscript.
  All conceptual contributions, theoretical choices, and interpretations remain
  the sole responsibility of the author.

  \bibliography{cosmochrony-bibliography,refs-external}

\end{document}
